Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals

### 2021-10-15 09_Article/01_Revision1_Reviewers/Simon_2021_Bone_Manuscript.tex
Finalize revision submission

	\begin{abstract}
		Osteogenesis Imperfecta (OI) is an inherited form of bone fragility characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass. High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology at peripheral sites including the weight-bearing distal tibia. The resulting 3D reconstructions can be used as a basis of micro-finite element (\si{\micro}FE) or homogenised finite element (hFE) models for bone strength estimation. The hFE scheme uses homogenised local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships. \Add{However, it is unclear if these relationships quantified previously for healthy controls are valid for trabecular bone from OI patients.} Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.

@@

		Compared to healthy controls, we found the OI samples to have significantly lower \Add{CV,} BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.). These results are in agreement with previous studies. \Del{The stiffness of ROIs from OI bone reached lower values compared to healthy controls and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range}. \Add{The stiffnesses of highly heterogeneous ROIs are randomly lower with respect to the fabric-elasticity relationships and reflects the limit of validity of the computational homogenisation methodology.}

@@

		In conclusion, \Add{OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and less dense than healthy controls.} Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis. \Del{Since highly heterogenous ROIs coincide with low BV/TV and can be handled smoothly by the fabric-elasticity relationships we expect them to play a minor role in hFE analysis of distal bone sections.} \Add{Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.} 
	\end{abstract}

@@

	\Add{OI occurs worldwide and without sex preference. OI type I is estimated to occur in 45\% of the total OI cases, type II about 10\%, type III 25\% and type IV 20\% \cite{Martin2007}}.Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels. Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018} \Add{but aBMD is limited for bone strength estimation \cite{Pasco2006}}.  \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well. Bone volume fraction (BV/TV) and trabecular number (Tb.N.) in OI bone is lower than for healthy controls. Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI compared to healthy control, and the trabecular thickness (Tb.Th.) is not significantly different. At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior (ultimate stress and ultimate strain) than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\

	High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}. Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008} \Add{used for fracture risk estimation}. \Add{Micro-finite element (\si{\micro}FE) analysis converts each voxel into an element and thus requires high performance computer for a bone section analysis}. Homogenized finite element (hFE) is based on \Add{local} BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}. High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}. \Add{Homogenised QCT-based FEA has become more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA).} Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk. However, HR-pQCT-based hFE relies on fabric-elasticity relationships. Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{ Similar fabric-elasticity relationships will allow to extend the application of HR-pQCT-based hFE of healthy bone to OI bone at least in the linear elastic regime.}

@@

	\subsection{HR-pQCT}
	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Br\"{u}tisellen,
	Switzerland) was performed at the distal tibia on all patients from both groups. \Add{The distal tibia was chosen because numerous studies were performed on the distal radius \cite{AriasMoreno2019,Boutroy2008,Gross2013,Hosseini2017,Kocijan2015,Varga2011}, some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) \cite{Boutroy2005,Folkestad2012} but only one study was performed on the distal tibia with XtremeCTII \cite{Schenk2020}.} Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}. Namely, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}. On the other hand, for member of the OI group, the reference line is placed at the subchondral endplate of the ankle joint.  The mid-point of the scanned volume of interest was located at a distance of 7\% of total tibia length from the reference line \cite{GHASEMZADEH2017,Bonaretti2017}, see Figure \ref{01_OI}. Then, the scans were acquired according to the manufacturer's standard in vivo acquisition protocol. The scanned regions according to these two different protocols coincide approximately and are shown side by side in Figure \ref{01_ClinicalSections}. For both the healthy and the OI groups, each stack consisted of 168 slices and an isotropic voxel resolution of 61 \si{\micro}m. This led to a thickness of roughly 10.2 mm for each stack. Standardized scanning settings were used (voltage of 60 kVp, 900 $\mu$A, 100 ms integration time) for the healthy group as well as for the OI group. For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale proposed by \citeauthor{Pialat2012} \cite{Pialat2012} of 1 (no motion artefacts) to 5 (extreme motion artefacts). The final grade of each scan was defined as the highest slice grade. For the OI group, as the scan consisted of one stack, only one grade was attributed using the same scale as for the control group. Scans were then processed independently of their quality grading. A summary of the scans grading is shown in Figure \ref{01_MotionArtefacts}.

@@

	\subsection{Image analysis}
	The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking)\Add{. Then, a Gauss filtering (sigma = 0.8, support = 1 voxel) was performed} and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}). Then, mask segmented images were used for further analysis.\\

	In general, six cubic ROIs were selected at random position in each scan in a predefined area. Each ROI had to contain trabecular bone, but no cortical bone. For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number \Del{3}\Add{1}) to be at roughly the same anatomical location as for the OI group. Then, for both the healthy group and the OI group, the stack was divided into two halves \Add{with a transverse plane} and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}. For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no enough trabecular bone. This led to 720 healthy ROIs from 120 individuals and 294 OI ROIs from 49 individuals.\\

@@

	A \si{\micro}FE analysis was performed using \textsc{ABAQUS 6.14}. In brief, each voxel of the cleaned ROI was converted to a fully integrated linear brick elements (C3D8) using a direct voxel conversion approach. Then, a Young's modulus $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element. The homogenisation process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}. Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system. Such rotation would potentially decrease image quality \Add{due to interpolation}. The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}. This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\

@@

	\subsection{Linear Regression}
	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model \Add{using the entire sample of both groups}. Standard linear models assume independent and identically distributed (iid) variables. As this assumption was violated by the fact that six ROIs were analysed per individual, a linear mixed-effect model was preferred. This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-independence of ROIs from the same individual. A more detailed form of this model is presented in Appendix \ref{A1}.

@@

	\section{Results}
	%
	\subsection{Morphological Analysis}
	The results of the morphological analysis are summarised in Table \ref{Table2}. The individual matching for age and sex allowed to have similar group distributions with 17 females and 11 males in each group. The mean age was 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched healthy and OI individuals, respectively. \Add{The OI group matched with control was composed of nineteen type I and nine type IV.} BV/TV of healthy individuals was higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01. Similarly, trabecular number was higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001. The trabecular thickness does not show significant differences between groups. On the other hand, permutation test performed for trabecular separation showed that it is higher in matched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$, -0.384] $\cup$ [0.421, $\infty$). Trabecular separation SD is higher in OI individuals compared to matched healthy individuals with a p-value of 0.02 and an exclusion range of ($-\infty$, -0.232] $\cup$ [0.251, $\infty$). SMI as well as the degree of anisotropy are higher for matched OI than for healthy people with p-values <0.001 and of 0.02, respectively. Finally, the log transformation of the coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [-0.757, -0.333] where CV is higher in matched OI individuals compared to matched healthy individuals.\\

@@

	\subsection{Comparison with literature}
	Table \ref{Table4} shows results obtained \Add{using HR-pQCT tibiae scans with a resolution of 61 \si{\micro}m} compared to \Del{literature} \Add{previous femora studies using \si{\micro}CT with a resolution of 18\si{\micro}m coarsened to 36\si{\micro}m \cite{Gross2013,Panyasantisuk2015}}. \citeauthor{Gross2013} \cite{Gross2013} has the larger number of ROIs. Data sets of \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} show BV/TV ranges slightly higher than in the present study and the one of \citeauthor{Gross2013} \cite{Gross2013}. On the other hand, DA is higher in the present study than for \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013}. Setting the exponents $k$ and $l$ to the same values led to lower stiffness constants for the observed data set compared to the other studies \cite{Gross2013,Panyasantisuk2015}.\\

@@

	\section{Discussion}
	Osteogenesis imperfecta is an inherited form of bone fra\-gility with a severity going from mild to perinatally lethal. This study aims to confirm that fabric-elasticity relationships in OI trabecular bone are similar compared to healthy bone, encouraging the use of HR-pQCT based hFE outcomes for fracture risk assessment. To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively. \Add{These numbers of participants are higher than in other studies comparing OI with healthy controls \cite{Folkestad2012,Kocijan2015,Rolvien2018}}.\\

@@

	After BV/TV and DA matching, grouping the data sets together led to similar $R^2_{adj}$ and NE as for the individual filtered data sets. This allows one to determine values for $k$ and $l$ for the tibia at a spatial resolution of 61 \si{\micro}m. Imposing these values to perform the linear regression on data sets of the matched individuals allows us to highlight differences, if any, between healthy and OI trabecular bone. The relatively low differences for $\lambda_0'$ and $\mu_0$ once again supports the hypothesis for similar fabric-elasticity relationships between healthy and OI trabecular bone. For $\lambda_0$, this relative difference being higher could rise some doubts about this similarity, but the 95\% CI intervals still show a common range which almost include both the $\lambda_0$ of OI and healthy linear regressions. Moreover, ANCOVA performed comparing the original formulation and the one with addition of a regressor for the group showed a p-value far above the 5\% significance level. With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone. \Add{Thus, the same hFE analysis can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones. Although} in hFE analysis \Del{of bones,} it is not possible to exclude part of the mesh because of high heterogeneity \Del{. Nevertheless}, \Add{the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.}\\

	Imposing $k$ and $l$ allows one to compare the other regressions parameters and estimate the effect of different image resolutions.  \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used femur scans with 18 \si{\micro}m spatial resolution and coarsened them to 36 \si{\micro}m. \citeauthor{Gross2013}\cite{Gross2013} showed that different anatomical locations lead to only slight differences. Comparing linear regression of the filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the higher DA range and by the coarser resolution. Differences of $R^2_{adj}$ and NE come from the imposition of $k$ and $l$ to a different value than the optimal ones. Then, comparing linear regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges overlap. As for the filtered data sets, DA is higher in the present study and the stiffness constants remain lower than for the two other studies. Here, differences in DA can mainly be explained by the different anatomical location and differences in stiffness constants as a result of the different image resolutions. The distal tibia, unlike the proximal femur, is mainly loaded in one direction which explains this higher DA. Lower stiffness constants are obtained because the coarser structure resulting from XCTII cannot be as optimised as the fine detailed structure obtained by \si{\micro}CT for the same BV/TV. Effectively, the architecture resulting from \si{\micro}CT scans can reproduce the optimised morphology of trabecular bone with higher fidelity. By decreasing the scan spatial resolution, the scanned structure becomes bulkier. Performing a linear regression on this less optimised structure leads to the observed lower stiffness constants. Finally, the comparison between $R^2_{adj}$ and NE of current study without imposing $k$ and $l$ and the ones of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and  \citeauthor{Gross2013}\cite{Gross2013} shows that lower spatial resolution leads to lower fit quality. Nevertheless, $l$ stays in the same range as for the two other studies \cite{Gross2013,Panyasantisuk2015}, meaning the relative weight of DA remains constant. On the other hand, the higher $k$ highlights an increased relative weight of BV/TV. \\

>>>>

	\begin{abstract}
		Osteogenesis Imperfecta (OI) is an inherited form of bone fragility characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass. High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology at peripheral sites including the weight-bearing distal tibia. The resulting 3D reconstructions can be used as a basis of micro-finite element (\si{\micro}FE) or homogenised finite element (hFE) models for bone strength estimation. The hFE scheme uses homogenised local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships. However, it is unclear if these relationships quantified previously for healthy controls are valid for trabecular bone from OI patients. Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.

@@

		Compared to healthy controls, we found the OI samples to have significantly lower CV, BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.). These results are in agreement with previous studies. The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology.

@@

		In conclusion, OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and less dense than healthy controls. Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.  Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.
	\end{abstract}

@@

	OI occurs worldwide and without sex preference. OI type I is estimated to occur in 45\% of the total OI cases, type II about 10\%, type III 25\% and type IV 20\% \cite{Martin2007}.Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels. Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018} but aBMD is limited for bone strength estimation \cite{Pasco2006}.  \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well. Bone volume fraction (BV/TV) and trabecular number (Tb.N.) in OI bone is lower than for healthy controls. Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI compared to healthy control, and the trabecular thickness (Tb.Th.) is not significantly different. At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior (ultimate stress and ultimate strain) than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\

	High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}. Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008} used for fracture risk estimation. Micro-finite element (\si{\micro}FE) analysis converts each voxel into an element and thus requires high performance computer for a bone section analysis. Homogenized finite element (hFE) is based on local BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}. High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}. Homogenised QCT-based FEA has become more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA). Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk. However, HR-pQCT-based hFE relies on fabric-elasticity relationships. Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships. Similar fabric-elasticity relationships will allow to extend the application of HR-pQCT-based hFE of healthy bone to OI bone at least in the linear elastic regime.

@@

	\subsection{HR-pQCT}
	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Br\"{u}tisellen,
	Switzerland) was performed at the distal tibia on all patients from both groups. The distal tibia was chosen because numerous studies were performed on the distal radius \cite{AriasMoreno2019,Boutroy2008,Gross2013,Hosseini2017,Kocijan2015,Varga2011}, some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) \cite{Boutroy2005,Folkestad2012} but only one study was performed on the distal tibia with XtremeCTII \cite{Schenk2020}. Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}. Name\-ly, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}. On the other hand, for member of the OI group, the reference line is placed at the subchondral endplate of the ankle joint.  The mid-point of the scanned volume of interest was located at a distance of 7\% of total tibia length from the reference line \cite{GHASEMZADEH2017,Bonaretti2017}, see Figure \ref{01_OI}. Then, the scans were acquired according to the manufacturer's standard in vivo acquisition protocol. The scanned regions according to these two different protocols coincide approximately and are shown side by side in Figure \ref{01_ClinicalSections}. For both the healthy and the OI groups, each stack consisted of 168 slices and an isotropic voxel resolution of 61 \si{\micro}m. This led to a thickness of roughly 10.2 mm for each stack. Standardized scanning settings were used (voltage of 60 kVp, 900 $\mu$A, 100 ms integration time) for the healthy group as well as for the OI group. For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale proposed by \citeauthor{Pialat2012} \cite{Pialat2012} of 1 (no motion artefacts) to 5 (extreme motion artefacts). The final grade of each scan was defined as the highest slice grade. For the OI group, as the scan consisted of one stack, only one grade was attributed using the same scale as for the control group. Scans were then processed independently of their quality grading. A summary of the scans grading is shown in Figure \ref{01_MotionArtefacts}.

@@

	\subsection{Image analysis}
	The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking). Then, a Gauss filtering (sigma = 0.8, support = 1 voxel) was performed and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}). Then, mask segmented images were used for further analysis.\\

	In general, six cubic ROIs were selected at random position in each scan in a predefined area. Each ROI had to contain trabecular bone, but no cortical bone. For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number 1) to be at roughly the same anatomical location as for the OI group. Then, for both the healthy group and the OI group, the stack was divided into two halves with a transverse plane and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}. For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no enough trabecular bone. This led to 720 healthy ROIs from 120 individuals and 294 OI ROIs from 49 individuals.\\

@@

	A \si{\micro}FE analysis was performed using \textsc{ABAQUS 6.14}. In brief, each voxel of the cleaned ROI was converted to a fully integrated linear brick elements (C3D8) using a direct voxel conversion approach. Then, a Young's modulus $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element. The homogenisation process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}. Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system. Such rotation would potentially decrease image quality due to interpolation. The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}. This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\

@@

	\subsection{Linear Regression}
	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model using the entire sample of both groups. Standard linear models assume independent and identically distributed (iid) variables. As this assumption was violated by the fact that six ROIs were analysed per individual, a linear mixed-effect model was preferred. This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-independence of ROIs from the same individual. A more detailed form of this model is presented in Appendix \ref{A1}.

@@

	\section{Results}
	%
	\subsection{Morphological Analysis}
	The results of the morphological analysis are summarised in Table \ref{Table2}. The individual matching for age and sex allowed to have similar group distributions with 17 females and 11 males in each group. The mean age was 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched healthy and OI individuals, respectively. The OI group matched with control was composed of nineteen type I and nine type IV. BV/TV of healthy individuals was higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01. Similarly, trabecular number was higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001. The trabecular thickness does not show significant differences between groups. On the other hand, permutation test performed for trabecular separation showed that it is higher in matched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$, -0.384] $\cup$ [0.421, $\infty$). Trabecular separation SD is higher in OI individuals compared to matched healthy individuals with a p-value of 0.02 and an exclusion range of ($-\infty$, -0.232] $\cup$ [0.251, $\infty$). SMI as well as the degree of anisotropy are higher for matched OI than for healthy people with p-values <0.001 and of 0.02, respectively. Finally, the log transformation of the coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [-0.757, -0.333] where CV is higher in matched OI individuals compared to matched healthy individuals.\\

@@

	\subsection{Comparison with literature}
	Table \ref{Table4} shows results obtained using HR-pQCT tibiae scans with a resolution of 61 \si{\micro}m compared to previous femora studies using \si{\micro}CT with a resolution of 18\si{\micro}m coarsened to 36\si{\micro}m \cite{Gross2013,Panyasantisuk2015}. \citeauthor{Gross2013} \cite{Gross2013} has the larger number of ROIs. Data sets of \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} show BV/TV ranges slightly higher than in the present study and the one of \citeauthor{Gross2013} \cite{Gross2013}. On the other hand, DA is higher in the present study than for \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013}. Setting the exponents $k$ and $l$ to the same values led to lower stiffness constants for the observed data set compared to the other studies \cite{Gross2013,Panyasantisuk2015}.\\

@@

	\section{Discussion}
	Osteogenesis imperfecta is an inherited form of bone fra\-gility with a severity going from mild to perinatally lethal. This study aims to confirm that fabric-elasticity relationships in OI trabecular bone are similar compared to healthy bone, encouraging the use of HR-pQCT based hFE outcomes for fracture risk assessment. To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively. These numbers of participants are higher than in other studies comparing OI with healthy controls \cite{Folkestad2012,Kocijan2015,Rolvien2018}.\\

@@

	After BV/TV and DA matching, grouping the data sets together led to similar $R^2_{adj}$ and NE as for the individual filtered data sets. This allows one to determine values for $k$ and $l$ for the tibia at a spatial resolution of 61 \si{\micro}m. Imposing these values to perform the linear regression on data sets of the matched individuals allows us to highlight differences, if any, between healthy and OI trabecular bone. The relatively low differences for $\lambda_0'$ and $\mu_0$ once again supports the hypothesis for similar fabric-elasticity relationships between healthy and OI trabecular bone. For $\lambda_0$, this relative difference being higher could rise some doubts about this similarity, but the 95\% CI intervals still show a common range which almost include both the $\lambda_0$ of OI and healthy linear regressions. Moreover, ANCOVA performed comparing the original formulation and the one with addition of a regressor for the group showed a p-value far above the 5\% significance level. With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone. Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones. Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.\\

	Imposing $k$ and $l$ allows one to compare the other regressions parameters and estimate the effect of different image resolutions.  \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used femur scans with 18 \si{\micro}m spatial resolution and coarsened them to 36 \si{\micro}m. \citeauthor{Gross2013}\cite{Gross2013} showed that different anatomical locations lead to only slight differences. Comparing linear regression of the filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the higher DA range and by the coarser resolution. Differences of $R^2_{adj}$ and NE come from the imposition of $k$ and $l$ to a different value than the optimal ones. Then, comparing linear regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges overlap. As for the filtered data sets, DA is higher in the present study and the stiffness constants remain lower than for the two other studies. Here, differences in DA can mainly be explained by the different anatomical location and differences in stiffness constants as a result of the different image resolutions. The distal tibia, unlike the proximal femur, is mainly loaded in one direction which explains this higher DA. Lower stiffness constants are obtained because the coarser structure resulting from XCTII cannot be as optimised as the fine detailed structure obtained by \si{\micro}CT for the same BV/TV. Effectively, the architecture resulting from \si{\micro}CT scans can reproduce the optimised morphology of trabecular bone with higher fidelity. By decreasing the scan spatial resolution, the scanned structure becomes bulkier. Performing a linear regression on this less optimised structure leads to the observed lower stiffness constants. Finally, the comparison between $R^2_{adj}$ and NE of current study without imposing $k$ and $l$ and the ones of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and  \citeauthor{Gross2013}\cite{Gross2013} shows that lower spatial resolution leads to lower fit quality. Nevertheless, $l$ stays in the same range as for the two other studies \cite{Gross2013,Panyasantisuk2015}, meaning the relative weight of DA remains constant. On the other hand, the higher $k$ highlights an increased relative weight of BV/TV.\\

	The analysis of TMD gives interesting outputs as well. The t-test reveals a 4\% higher TMD in OI trabecular bone than in healthy condition, agreeing with \citeauthor{Boyde1999}\cite{Boyde1999}, with 95\% certainty and a very high significance level, despite similar BV/TV and DA. The linear regression shows a significant relation between TMD and BV/TV, that is parallel in the healthy and OI groups (Figure \ref{02_TMD}). This can have two different origins. From a biological point of view, a mineralisation gradient develops between the core of a trabecula and the outer surface. The core undergoes secondary mineralisation while the surface is subjected to remodelling and deposition of younger, less mineralised matrix. An increase in BV/TV reduces the relative proportion of bone surface bone per unit bone volume (BS/BV), which would therefore raise the mean TMD. The second explanation comes from the partial volume effect of the scanner that reduces the intensity of surface voxels. Following the same argument, the reduced ratio of bone surface to bone volume (BS/BV) raises the mean TMD with increasing BV/TV. The former (biological) explanation is expected to have a lower impact than the latter (scanning). Nevertheless, the intercept of the linear regression depends on the group i.e. TMD is significantly higher in the OI group over the entire BV/TV range. This could be explained by the biphosphonate treatment that OI patients receive. Indeed, bisphosphonates slow down the remodelling process which lead to higher (secondary) mineralization of the bone. According to the recent results of \citeauthor{Indermaur2021}\cite{Indermaur2021}, our findings suggest adding a correction for TMD in FE simulations to account for the resulting increase in modulus, ultimate stress and post-yield properties of OI bone compared to healthy bone at the ECM level.\\

	The main limitations of this study are the definition of "homogeneous enough" (i.e. low CV) and the fact that it is limited to XCTII reconstructions of the distal tibia. Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types. ROI homogeneity, more specifically a balanced bone connectivity of the 6 faces of the ROI, is a necessary condition underlying the computational homogenisation of mechanical properties and the results of heterogenous samples must be disregarded. However, when calibrated with homogeneous ROIs, fabric-elas\-ticity relationships offer a smooth and reasonable extrapolation of stiffness properties for the concerned low-homogeneity and low-stiffness ROIs. In fact, these low-property ROIs were not shown to compromise the performance of hFE models of the distal radius \cite{Varga2011, Hosseini2017, AriasMoreno2019}.\\
